---
title: "CarrefourAnomalyDetection"
author: "Joy Machuka"
date: "9/11/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Definition
As a Data analyst at Carrefour Kenya working for the  marketing department Carrefour you are to come up with  relevant marketing strategies that will result in the higher no. of sales. Amongst the processes is checking for anomalies in the data.

# Specifying the Question
Identify Anomalies in the Sales of Carrefour

# Metric of Success
When we are able to detect anomalies in the sales trend.

#Context
Anomalies could be used in sales data to detect fraud or to detect abnormal trends in sales made hence further investigation. In our case we are going to check for both.

# Experimental Design
Load the Data
Check the Data
Data cleaning
Exploratory Data Analysis
Implement the Solution
Challenge the Solution

# Data Sourcing
```{r}
# loaddata
sales <- read.csv("http://bit.ly/CarreFourSalesDataset")
head(sales)
```

```{r}
dim(sales)
```
Data has 1000 columns and 2 columns
```{r}
class(sales)
```
```{r}
str(sales)
```
Our columns are  date and sales. 
```{r}
#checking for missing values
anyNA(sales)
```
No missing values
```{r}
library(tidyverse)
```
```{r}
library(tibbletime)
```
```{r}
sales$Date <- as.Date(sales$Date, "%m/%d/%Y")

sales_ts = sales %>%
  as_tibble() %>%
  as_tbl_time(Date) %>%
  arrange(Date) %>%
  as_period("daily")
```

```{r}
anyNA(sales_ts)
```
Implementing solution
```{r}
# install.packages("anomalize")
library(anomalize)
```

```{r}
library(dplyr)
```

```{r}
pkg <- c('timetk')
# install.packages(pkg)
```
```{r}
library(timetk)
```

```{r}
#function
sales_anomaly <- sales_ts %>%
  time_decompose(Sales) %>% 
  anomalize(remainder,max_anoms = 0.2, alpha=0.05) %>%
  time_recompose() %>% glimpse()

```

```{r}
# Plot anomalies
sales_anomaly %>% plot_anomalies(time_recomposed = TRUE)

```
Our data has no anomalies.





