---
title: "CarrefourAssociationAnalysis"
author: "Joy Machuka"
date: "9/11/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Definition
As a Data analyst at Carrefour Kenya working for the  marketing department Carrefour you are to come up with  relevant marketing strategies that will result in the higher no. of sales. Amongst the processes is checking for associations in the data.

# Specifying the Question

Identify relationships between the Group of Items

# Metric of Success
Identify the relationships between the Groups of Items provided

# Context
Working on the Carrefour Kenya Sales data to identify the most relevant marketing strategies by identifying related groups of Items.

# Experimental Design
Load the Data
Check the Data
Implement the Solution
Challenge the Solution

# Data Sourcing
```{r}
library(arules)
```

```{r}
# Data to find associations
association <- read.transactions("http://bit.ly/SupermarketDatasetII")
head(association)
```
```{r}
dim(association)
```
```{r}
class(association)
```

```{r}
#check first 10 items
inspect(association[1:10])
```
```{r}
# number of items in the first 10 itemset or transaction
size(head(association, 10))
```
```{r}
item <- as.data.frame(itemLabels(association))
colnames(item) <- "Item"
head(item,10)
```
```{r}
summary(association)
```
```{r}
print("Frequency of each item")
itemFrequency(association[,1:15], type="absolute")
cat ("Tab\n")

print("Frequency of first 15 Items compared to Others")
# How many times these items have appeared as compared to others.
round(itemFrequency(association[,1:15], type="relative")*100,2)
```
```{r}
# Most common items
library(RColorBrewer)
itemFrequencyPlot(association, topN= 10, col=brewer.pal(8,'Set1'),main="Relative Item Frequency Plot")
itemFrequencyPlot(association, support= 0.1, col=brewer.pal(8,'Set1'),main="Relative Importance is at least 10%")
```
```{r}
# Build a Model based on association rules
# min_supp -> 0.001
# confidence -> 0.8
rules <- apriori(association, parameter = list(supp=0.001, conf=0.75))
rules
summary(rules)
inspect(rules[1:5])
```
```{r}
rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:10])
```

The  5 most popular Items are:-

1. tea
2. wheat
3. mineal
4. fat
6. yoghurt

The 5 most popular have support at least 10% and 2 more items which are:-

1. green tea 
2. milk

We have 4 rules with 100% confidence in the association. We can use this association to increase sales in Females customers in the Branch A and B buy identifying which products the Female Customers buy most
This can also be done for Male Customers in Branch C


The Model can also be used to improve the Marketing and Sales of the Product Lines that are not popular with any of the customers in any Branch.


