---
title: "CarrefourMarketingAnalysis"
author: "Joy Machuka"
date: "9/10/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Statement
Carrefour needs to increase their sales. They therefore get a data analyst to help them come up with marketing strategies to help them achieve this.

# Defining the Question
You are a Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).

# Metrics of Success
Our analysis will be considered successful when we are able to perform an efficient dimensionality reduction and feature selection and come up with good marketing strategies thereafter.

# Context
Carrefour is one of the supermarkets in Kenya. They are ranked amongst the top shopping centres in Kenya hence they are always seeking to offer good customer service and in turn employ effective marketing strategies to increase their sales.

# Experimental Design
Load the Data
Checking the Data
Data cleaning
Exploratory Data Analysis
Implement the Solution
Challenge the Solution

# Loading the data
```{r}
carrefour <- read.csv("http://bit.ly/CarreFourDataset")
head(carrefour)
```
```{r}
class(carrefour)
```
```{r}
dim(carrefour)
```
Our dataframe has 1000 rows and 16 columns.
```{r}
names(carrefour)
```
```{r}
str(carrefour)
```
We have 8 numerical columns and 8 char columns. We will convert the char categorical columns to factors for analysis.

# Data Cleaning
```{r}
#checking for nulls
anyNA(carrefour)
```
We have no missing values.
```{r}
#checking for duplicates
duplicated_rows <- carrefour[duplicated(carrefour),]
duplicated_rows
```
We have no duplicate values
```{r}
#change column names to lower case
lower <- function(x) {
    colnames(x) <- tolower(colnames(x))
    x
}
carrefour <- lower(carrefour)
names(carrefour)
```
```{r}
#drop the ID column
carrefour1 <- carrefour[,-1]
names(carrefour1)
```
```{r}
dim(carrefour1)
```
New dataframe has now 15 columns 1000 rows
```{r}
str(carrefour1)
```

```{r}
num_cols <- carrefour1[, c(5,6,7,11,13,14,15)]
names(num_cols)
```

```{r}
outliers = function(x){
  for(i in colnames(x)){
    boxplot(carrefour1[[i]], xlab=i, main=paste0("Boxplot for ",i))
  }
}
outliers(num_cols)
```
```{r}
#replacing outliers with 5th and 95th percentile
outreplace <- function(x){
   qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
   caps <- quantile(x, probs=c(.05, .95), na.rm = T)
   H <- 1.5 * IQR(x, na.rm = T)
   x[x < (qnt[1] - H)] <- caps[1]
   x[x > (qnt[2] + H)] <- caps[2]
   return(x)
}
carrefour1$tax <- outreplace(carrefour1$tax)
carrefour1$cogs <-outreplace(carrefour1$cogs)
carrefour1$gross.income <- outreplace(carrefour1$gross.income)
carrefour1$total <- outreplace(carrefour1$total )
```

```{r}
outliers = function(x){
  for(i in colnames(x)){
    boxplot(carrefour1[[i]], xlab=i, main=paste0("Boxplot for ",i))
  }
}
outliers(num_cols)
```
We have removed all outliers from the numerical columns

#Exploratory Data Analysis
##Univariate Analysis
```{r}
library(dplyr)
carrefour1 <- mutate_at(carrefour1, vars(branch, customer.type,gender,product.line,payment), as.factor)
str(carrefour1)
```

```{r}
#Summary statistics
summary(carrefour1)
```
branch A has 340, branch 332 and branch C has 328 transactions.
Females are 501 and 499 males
344 cash payment 311 credit cat payments and 345 using Ewallet
There were 170 electronic accessories, 178 fashion accessories,174 food and beverages,152 health and beauty, 160 home and lifestyle and 166 sports and travel.
```{r}
#mode function
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
#apply it on the duration columns
getmode(carrefour1$quantity)
```
Most number of purchased items per invoice were 10
```{r}
#descriptive statistics of the dataframe
psych::describe(carrefour1)
```
```{r}
#plotting histograms of the numerical columns
histogram = function(x){
  for(i in colnames(x)){
    hist(carrefour1[[i]], breaks = 10,main =i,xlab = i,col = "lightblue")
  }
}
histogram(num_cols)
```
```{r}
library(ggplot2)
# Customer Type
customerdist <- ggplot(carrefour1 ,aes(x=customer.type , fill=customer.type)) + geom_bar() + labs(title = "Distribution of the Customer Types", x="Customer Type")
customerdist +scale_fill_discrete(name = "Customer Type")

# How many Branches we have
branchdist <- ggplot(carrefour1, aes(x=branch, fill=branch)) + geom_bar()+labs(title = "Distribution of Branches", x="Branches")+scale_fill_discrete(name = "Branch") 
branchdist

# Gender Distribution
genderdist <- ggplot(carrefour1, aes(x=gender, fill=gender)) + geom_bar()+labs(title = "Distribution of Gender", x="Gender")+scale_fill_discrete(name = "Gender") 
genderdist

# Types Product Line
productlinedist <- ggplot(carrefour1,aes(x=product.line, fill=product.line))+ geom_bar()+ labs(title = "Distribution of Product Line", x="Product Line")+ scale_fill_discrete(name = "Product Line")
productlinedist + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# Types Payments
paymentdist <- ggplot(carrefour1, aes(x=payment, fill=payment)) + geom_bar()+labs(title = "Distribution of Payments", x="Payment")+scale_fill_discrete(name = "Payment") 
paymentdist
```
## Bivariate Analysis
```{r}
#Customer Type groupedby Branch
c_typecomparison <- ggplot(carrefour1, aes(x=branch, fill=customer.type)) + geom_bar(position = "dodge")+labs(title = "Distribution of Customer Type per Branch", x="Payment")+scale_fill_discrete(name = " Customer Type") 
c_typecomparison

#Gender per Branch
gendercomp <- ggplot(carrefour1, aes(x=branch, fill=gender)) + geom_bar(position = "dodge")+labs(title = "Distribution of Gender per Branch", x="Branch")+scale_fill_discrete(name = "Gender") 
gendercomp

#Payment mode per Branch
paymentcomp <- ggplot(carrefour1, aes(x=branch, fill=payment)) + geom_bar(position = "dodge")+labs(title = "Distribution of Payment Mode per Branch", x="Branch")+scale_fill_discrete(name = "Payment") 
paymentcomp

# Customer Type vs Mode of Payment
c_typecomp2 <- ggplot(carrefour1, aes(x=customer.type, fill=payment)) + geom_bar(position = "dodge")+labs(title = "Customer Type and Mode of Payment", x="Customer Type")+scale_fill_discrete(name = "Mode of Payment") 
c_typecomp2

# Customer Type and Product Line
prodlinecomp2 <- ggplot(carrefour1, aes(x=product.line, fill=customer.type)) + geom_bar(position = "dodge")+labs(title = "Distribution of Product Line for Each Customer Type", x="Product Line")+scale_fill_discrete(name = "Customer Type") 
prodlinecomp2 +theme(axis.text.x = element_text(angle = 30, hjust=0.8))

# Product Line vs Gender
prodlinecomp3 <- ggplot(carrefour1, aes(x=product.line, fill=gender)) + geom_bar(position = "dodge")+labs(title = "Distribution of Product Line in Each Gender", x="Product Line")+scale_fill_discrete(name = "Gender") 
prodlinecomp3 +theme(axis.text.x = element_text(angle = 30, hjust=0.8))

#Product line vs quantity per branch
prodlinecomparison <- ggplot(carrefour1,aes(x=product.line,y=quantity,fill=product.line))+geom_bar(stat = "identity")+ facet_wrap(~branch, scales = "free") + labs(title = "Total Quantity of each Product Line", subtitle = "[per Branch]",x="Product Line", y="Total Quantity") + scale_fill_discrete(name = "Product Line") + theme(axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank())
prodlinecomparison
```
```{r}
# Product Line with Highest Gross Income
plineby_gross <- carrefour1 %>%
  group_by(product.line) %>%
  summarise(gross = sum(gross.income))
gross_inc <- ggplot(plineby_gross, aes(x=product.line, y= gross, fill=product.line))+geom_bar(stat = "identity")+ labs(title = "Product Line vs Gross Income", x="Product Line",y="Total Gross Income")
gross_inc + scale_fill_brewer(name="Product Line",palette="Set1")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```
```{r}
# Product Line total sales
plineby_sale <- carrefour1 %>%
  group_by(product.line) %>%
  summarise(total = sum(total))
total_sale <- ggplot(plineby_sale, aes(x=product.line, y= total, fill=product.line))+geom_bar(stat = "identity")+ labs(title = "Total Sales from Each Product Line", x="Product Line",y="Total Sales")
total_sale + scale_fill_brewer(name="Product Line",palette="Set1")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```
```{r}
#Banch ratings
branchby_rate <- carrefour1 %>%
  group_by(branch) %>%
  summarise(rate = mean(rating))
rate <- ggplot(branchby_rate, aes(x=branch, y= rate, fill=branch))+geom_bar(stat = "identity")+ labs(title = "Average Rating Per Branch", x="Branch",y="Average Rating")
rate + scale_fill_brewer(palette="Set1")+theme(legend.position = "none") 
```
```{r}
#Branch totalsales
branchby_sale <- carrefour1 %>%
  group_by(branch) %>%
  summarise(total = sum(total))
total_sale <- ggplot(branchby_sale, aes(x=branch, y= total, fill=branch))+geom_bar(stat = "identity")+ labs(title = "Total Sales from Each Branch", x="Branch",y="Total Sales")
total_sale + scale_fill_brewer(name="Branch",palette="Set1")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```
```{r}
#customer type  gross income comparison
custtypeby_gross <- carrefour1 %>%
  group_by(customer.type) %>%
  summarise(gross = sum(gross.income))
cust_type <- ggplot(custtypeby_gross, aes(x=customer.type, y= gross, fill=customer.type))+geom_bar(stat = "identity")+ labs(title = " Total Gross Income per Customer Type", x="Customer Type",y="Total Gross Income")
cust_type + scale_fill_brewer( name= "Customer Type",palette="Set1")+theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```
```{r}
# Covariance
covariance = cov(num_cols)
View(round(covariance,2))
```

```{r}
# Correlation matrix
corr_matrix = cor(num_cols)
corr <- as.data.frame(round(corr_matrix,2))
corr
```
# PCA
```{r}
str(carrefour1)
```
```{r}
num_features <- carrefour1[, c(5,6,7, 11,13:15)]
names(num_features)
```
```{r}
carrefour.pca <- prcomp(num_features, center = TRUE, scale. = TRUE)
carrefour.pca
summary(carrefour.pca)
```
```{r}
str(carrefour.pca)
```
# Feature Selection

```{r}
library(caret)
library(corrplot)
```

```{r}
highlyCorrelated <- findCorrelation(corr_matrix, cutoff=0.75)
```

```{r}
highlyCorrelated

names(carrefour1[,highlyCorrelated])
```
```{r}
carrefour2<-carrefour1[-highlyCorrelated]
str(carrefour2)
```
```{r}
num_features2 <- carrefour2[, c(3,4,8, 10:12)]
names(num_features2)
```

```{r}
# Performing our graphical comparison
# ---
# 
par(mfrow = c(1, 2))
corrplot(corr_matrix, order = "hclust")
corrplot(cor(num_features2), order = "hclust")
```
Variables acceptable were: unit.price, quantity, cogs, gross.income, rating and total"

Wrapper Method
```{r}
library(clustvarsel)
library(mclust)
```

```{r}
head(num_features2)
```
```{r}
# out = clustvarsel(num_features2, G=1:6)
# out
# # Features selected
# data <- num_features2[,out$subset]
# head(data)
```
Our wrapper method accepted gross income but was taking long to complete the search hence terminated it.

# Feature ranking
```{r}
library(FSelector)

feat <- num_features2
Scores <- linear.correlation(total~., feat)
Scores
# Selected Features
Subset <- cutoff.k(Scores, 5)
as.data.frame(Subset)
```
With feature ranking we had cogs, gross.income, quantity, unit.price and rating








